---
title: "Martensite Starting Temperature"
author: "Hooman Sabarou & Mounika Chevva-Cross_Validation Group_Fall 2024"
editor: visual
format: html
self-constrainded: true
echo: FALSE
---

### The head of the dataset

```{r}
# Load the dataset
dataset <- read.csv("/Users/hoomansabarou/Downloads/Martensite Start-Temp.csv")
head(dataset)

```

### Checking for missing values

```{r}
# Check for missing values
sum(is.na(dataset))

```

### Checking the structure of the dataset

```{r}
# Check the structure of the dataset
str(dataset)

```

### Summary statistics of the dataset

```{r}
# Summary statistics of the dataset
summary(dataset)

```

### Correlation between variables

```{r}
# Correlation between variables
cor_matrix <- cor(dataset)
cor_matrix

```

### Scatter plot for one element and Martensite starting temperature

```{r}
# Scatter plot for one element vs Ms
plot(dataset$C, dataset$Ms, xlab = "C", ylab = "Martensite Start Temp")

```

### Histogram of Martensite starting temperature

```{r}
# Plot histogram of Ms (Martensite start temperature)
hist(dataset$Ms, 
     main = "Distribution of Martensite Start Temperature (Ms)", 
     xlab = "Martensite Start Temperature (Ms)", 
     col = "lightblue", 
     border = "black", 
     breaks = 30)

```

### Filtering the dataset for temperatures between 300 to 800 degree Celcius

```{r}
# Filter the dataset for Ms between 300°C and 800°C
filtered_dataset <- subset(dataset, Ms >= 300 & Ms <= 800)

```

### Building GLM models

```{r}
# Build the updated GLM model using filtered data
updated_model <- glm(Ms ~ C + Ni + Mn + Mo + Si + Cr + V + Co, data = filtered_dataset, family = gaussian)

# Summarize the updated model
summary(updated_model)

```

### Checking residuals

```{r}
# Check residuals with a QQ plot
qqnorm(residuals(updated_model))
qqline(residuals(updated_model), col = "red")

# Histogram of residuals
hist(residuals(updated_model), main = "Residuals Distribution", xlab = "Residuals")


```

### Identifying outliers

```{r}
# Identify outliers
outlier_residuals <- which(abs(residuals(updated_model)) > 2 * sd(residuals(updated_model)))
dataset[outlier_residuals, ]

```

### Remove rows where Carbon content is zero, update the model, and the summary of the model. 

```{r}
# Remove rows where C is 0
filtered_dataset_no_C0 <- subset(filtered_dataset, C != 0)

# Build the updated GLM model using the filtered data
updated_model_no_C0 <- glm(Ms ~ C + Ni + Mn + Mo + Si + Cr + V + Co, data = filtered_dataset_no_C0, family = gaussian)

# Summarize the updated model
summary(updated_model_no_C0)

```

### Updating the model by removing Mo and Co and adding an interaction parameter between Carbon and Manganese.

```{r}
# Update the GLM model by removing Co and Mo, and adding an interaction between C and Mn
updated_model_interaction <- glm(Ms ~ C * Mn + Ni + Si + Cr + V, 
                                 data = filtered_dataset_no_C0, family = gaussian)

# Summarize the updated model
summary(updated_model_interaction)

```

### Update the model by removing V and add an interaction parameter between Carbon and Nickel. 

```{r}
# Update the model: remove V, and add interaction between C and Ni
updated_model_C_Ni_interaction <- glm(Ms ~ C * Mn + C * Ni + Si + Cr, 
                                      data = filtered_dataset_no_C0, 
                                      family = gaussian)

# Summarize the updated model
summary(updated_model_C_Ni_interaction)

```

### Performing ANOVA for the updated model. 

```{r}
library(classpackage)
# Perform ANOVA on the updated model
anova_check(updated_model_C_Ni_interaction)

```

```{r}
library(glmtoolbox)
(adjR2(updated_model_C_Ni_interaction))
```

### Checking leverage and influential points.

```{r}
cooks(updated_model_C_Ni_interaction)
```

### Investigate the influential points

```{r}
# Investigate the influential points
influential_points <- c(1081, 685, 686)
filtered_dataset_no_C0[influential_points, ]


```

### Remove influential points

```{r}
# Remove the influential points
filtered_dataset_no_influential <- filtered_dataset_no_C0[-c(1081, 685, 686), ]

# Rebuild the model after removing the influential points
updated_model_no_influential <- glm(Ms ~ C * Mn + C * Ni + Si + Cr, 
                                    data = filtered_dataset_no_influential, 
                                    family = gaussian)

# Summarize the updated model
summary(updated_model_no_influential)

```

### Investigate points found out by Cook's distances

```{r}
# Investigate the observations with moderate Cook's Distance
moderate_influential_points <- c(1132, 1188, 1061)
filtered_dataset_no_C0[moderate_influential_points, ]

```

### Apply the residuals check to the filtered dataset

```{r}
# Apply the residuals check to the correct dataset (filtered_dataset_no_influential)
filtered_dataset_no_influential <- filtered_dataset_no_influential %>% 
  mutate(outlier = if_else(abs(rstandard(updated_model_no_influential)) > 2.5, 
                           "Suspected", "Not Suspected"))

```

### Counting outliers

```{r}

filtered_dataset_no_influential %>% count(outlier)
```

### Plotting Carbon content with Martensite starting temprature

```{r}
library(ggplot2)
filtered_dataset_no_influential %>% ggplot(aes(x = C, y = Ms, color = outlier)) +
  geom_point() + 
  scale_color_manual(values = c("#999999", "#000000")) +
  labs(x = "Carbon", y = "Martensite Starting Temperature", color = "Outlier") +
  theme_bw()
```

```{r}
anova_check(updated_model_no_influential)

```

### Check Multicollinearity

```{r}
#Check Multicollinearity
car::vif(updated_model_no_influential)

```

### Remove identified outliers

```{r}
#Removing identified outliers
filtered_dataset_no_influential_no_outlier <- filtered_dataset_no_influential %>% filter(outlier == "Not Suspected")

updated_model_no_influential_2 <- glm(Ms ~ C * Mn + C * Ni + Si + Cr, 
                                    data = filtered_dataset_no_influential_no_outlier, 
                                    family = gaussian)

summary(updated_model_no_influential_2)

```

```{r}
anova_check(updated_model_no_influential_2)
```

```{r}
cooks(updated_model_no_influential_2)
```

### Table 1

```{=html}
<style type="text/css">
.tg  {border-collapse:collapse;border-spacing:0;}
.tg td{border-color:black;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;
  overflow:hidden;padding:10px 5px;word-break:normal;}
.tg th{border-color:black;border-style:solid;border-width:1px;font-family:Arial, sans-serif;font-size:14px;
  font-weight:normal;overflow:hidden;padding:10px 5px;word-break:normal;}
.tg .tg-0lax{text-align:left;vertical-align:top}
</style>
```
|     |       Ms      |       C        |       Mn       |       Ni       |       Si       |       Cr       |
|-----------|-----------|-----------|-----------|-----------|-----------|-----------|
|     | Min.   :310.0 | Min.   :0.0016 | Min.   :0.0000 | Min.   : 0.000 | Min.   :0.0000 | Min.   : 0.000 |
|     | 1st Qu.:553.5 | 1st Qu.:0.1600 | 1st Qu.:0.4700 | 1st Qu.: 0.000 | 1st Qu.:0.2000 | 1st Qu.: 0.000 |
|     | Median :605.0 | Median :0.3300 | Median :0.6900 | Median : 0.150 | Median :0.2600 | Median : 0.520 |
|     | Mean   :601.8 | Mean   :0.3617 | Mean   :0.7917 | Mean   : 1.558 | Mean   :0.3475 | Mean : 1.043   |
|     | 3rd Qu.:670.0 | 3rd Qu.:0.4400 | 3rd Qu.:0.9700 | 3rd Qu.: 1.580 | 3rd Qu.:0.3400 | 3rd Qu.: 1.150 |
|     | Max.   :784.0 | Max.   :1.4600 | Max.   :4.9500 | Max.   :27.200 | Max.   :3.8000 | Max.   :16.200 |

```{r}
write.csv(summary(filtered_dataset_no_influential_no_outlier[c("Ms", "C", "Mn", "Ni", "Si", "Cr")]), "NewTable.csv")
```
